### 2024-06-12 supplements/supplements.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{unicode-math}
\usepackage{setspace}
\usepackage{libertine}
\captionsetup[figure]{font={stretch=1}}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Supplements for the manuscript `Measuring variation in gaze following across communities, ages, and individuals --- a showcase of the TANGO-CC'},
  pdfauthor={Julia Christin Prein, Florian M. Bednarski, Ardain Dzabatou, Michael C. Frank, Annette M. E. Henderson, Josefine Kalbitz, Patricia Kanngiesse, Dilara Keşşafoğlu, Bahar Köymen, Maira V. Manrique-Hernandez, Shirley Magazi, Lizbeth Mújica-Manrique, Julia Ohlendorf, Damilola Olaoba, Wesley R. Pieter, Sarah Pope-Caldwell, Umay Sen, Katie Slocombe, Robert Z. Sparks, Roman Stengelin, Jahnavi Sunderarajan, Kirsten Sutherland, Florence Tusiime, Wilson Vieira, Zhen Zhang, Yufei Zong, Daniel B. M. Haun, \& Manuel Bohn},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Supplements for the manuscript `Measuring variation in gaze following across communities, ages, and individuals --- a showcase of the TANGO-CC'}
\author{Julia Christin Prein\textsuperscript{}, Florian M. Bednarski\textsuperscript{}, Ardain Dzabatou\textsuperscript{}, Michael C. Frank\textsuperscript{}, Annette M. E. Henderson\textsuperscript{}, Josefine Kalbitz\textsuperscript{}, Patricia Kanngiesse\textsuperscript{}, Dilara Keşşafoğlu\textsuperscript{}, Bahar Köymen\textsuperscript{}, Maira V. Manrique-Hernandez\textsuperscript{}, Shirley Magazi\textsuperscript{}, Lizbeth Mújica-Manrique\textsuperscript{}, Julia Ohlendorf\textsuperscript{}, Damilola Olaoba\textsuperscript{}, Wesley R. Pieter\textsuperscript{}, Sarah Pope-Caldwell\textsuperscript{}, Umay Sen\textsuperscript{}, Katie Slocombe\textsuperscript{}, Robert Z. Sparks\textsuperscript{}, Roman Stengelin\textsuperscript{}, Jahnavi Sunderarajan\textsuperscript{}, Kirsten Sutherland\textsuperscript{}, Florence Tusiime\textsuperscript{}, Wilson Vieira\textsuperscript{}, Zhen Zhang\textsuperscript{}, Yufei Zong\textsuperscript{}, Daniel B. M. Haun\textsuperscript{}, \& Manuel Bohn\textsuperscript{}}
\date{}

\shorttitle{supplements}

\affiliation{\phantom{0}}

\begin{document}
\maketitle

\section{Voice-over script}\label{voice-over-script}

The TANGO--CC was designed for a cross-cultural study by Bohn et al. (2024). For further information on the sample and further methodological choices, see their supplementary material ().

Here, we repeat the voice over script as it was presented in Bohn et al. (2024). The first trial of each trial type played audio descriptions of the presented events, and the voice-over script in English was as follows:

\subsection{Welcome (before the start of the task)}\label{welcome-before-the-start-of-the-task}

Hello! Great that you're here. We'll now play a balloon game. Can you see the children in the picture over there? We want to play together with the children using the balloon. We'll now talk you through exactly what will happen.

\subsection{Training 1}\label{training-1}

Look, a girl/boy is standing in the window. And can you see the balloon over there? The balloon always falls down and lands on the ground. And you have to find it! The girl/boy helps you and always looks at the balloon.

{[}balloon falls down{]}

Where is the balloon? Touch the balloon!

\subsection{Training 2}\label{training-2}

Perfect, that was great! Now, we'll continue playing. Can you see the girl/boy and the balloon again? The balloon will fall down again. This time, it will fall behind a hedge. And you have to find it! The girl/boy helps you and looks at the balloon.

{[}balloon falls down{]}

Where is the balloon? On the hedge, touch where the balloon is.

\subsection{Test}\label{test}

Nice, good job! Now, we'll continue playing. There is the balloon, the girl/boy and the hedge. The hedge is growing a bit now.

The balloon is behind the hedge now. You can't see it - but the girl/boy can! The balloon falls to the ground and you have to find it. Remember - the girl/boy always looks at the balloon!

\subsection{Goodbye (after the last test trial)}\label{goodbye-after-the-last-test-trial}

The children are super happy after playing. Thanks a lot for your help! See you soon and goodbye!

\section{Default settings}\label{default-settings}

When researchers use the default settings of the TANGO-CC, the following settings are applied based on the selected language. Visit this website \url{https://ccp-odc.eva.mpg.de/tango-cc/customize.html} and hover over the agent faces to view their IDs.

\begin{itemize}
\tightlist
\item
  Bemba (Zambia): background 01, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Chinese (China): background 01, agents f18-f19-f20-f21-m18-m19-m20-m21
\item
  English (India): background 01, agents f22-f23-f24-f25-m22-m23-m24-m25
\item
  English (New Zealand): background 01, agents f05-f06-f11-f15-m05-m06-m11-m15
\item
  English (Nigeria): background 01, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  English (UK): background 01, agents f05-f07-f08-f10-m03-m06-m09-m11
\item
  English (USA): background 01, agents f01-f05-f06-f09-m05-m06-m08-m09
\item
  German (Germany): background 01, agents f05-f07-f08-f10-m03-m06-m09-m11
\item
  Hai\textbar\textbar kom (Namibia): background 01, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Khwedam (Namibia): background 01, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Lingala (Rep.~Congo): background 02, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Marathi (India): background 01, agents f22-f23-f24-f25-m22-m23-m24-m25
\item
  Shona (Zimbabwe): background 01, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Spanish (Argentina): background 01, agents f05-f07-f09-f21-m05-m07-m13-m18
\item
  Spanish (Mexico): background 01, agents f05-f12-f15-f16-m13-m14-m15-m17
\item
  Kiswahili (Uganda): background 02, agents f01-f02-f03-f04-m01-m02-m03-m04
\item
  Turkish (Türkiye): background 01, agents f05-f09-f18-f21-m05-m09-m18-m21
\item
  Yaka (Rep.~Congo): background 02, agents f01-f02-f03-f04-m01-m02-m03-m04
\end{itemize}

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-bohn2024universal}
Bohn, M., Prein, J. C., Ayikoru, A., Bednarski, F. M., Dzabatou, A., Frank, M. C., Henderson, A. M. E., Isabella, J., Kalbitz, J., Kanngiesser, P., Keşşafoğlu, D., Koymen, B., Manrique-Hernandez, M., Magazi, S., Mújica-Manrique, L., Ohlendorf, J., Olaoba, D., Pieters, W., Pope-Caldwell, S., \ldots{} Haun, D. (2024). \emph{A universal of human social cognition: {Children} from 17 communities process gaze in similar ways}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/z3ahv}

\end{CSLReferences}

\endgroup

\end{document}