### 2024-11-26 stats/analysis.Rmd
add README files for project and data explanation, add comments to analysis script, update figures for production

---
title: "Tango-cc analysis"
author: "Julia Prein"
date: "11/26/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
# Load packages
library(tidyverse) # data handling etc.
library(tidybayes) # for stat_halfeye, add_predictive_draws
library(tidyboot) # for bootstrapped CIs
library(brms) # modeling
# library(cmdstanr) # modeling
library(ggpubr) # ggarrange
library(splithalfr) # split half reliability
library(ggridges) # for ridge plot
# load source code for geom_flat_violin()
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")

options(scipen = 999) # turn off scientific notation
theme_set(theme_light()) # set theme

# Seed for random number generation
set.seed(42)
```

```{r load_data}
# Read in data set 
complete_data <- read_csv("../data/data.csv")
```

```{r data_preparation}
# Prepare data set, clean community names, create continent variable, calculate imprecision, centrality, age group
# complete dataset, with all trials
all_trials <- complete_data %>% 
    mutate(
      continent = case_when(community %in% c("leipzig", "plymouth") ~ "Europe", 
                            community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda") ~ "Africa",
                            community %in% c("stanford", "mexico", "buenos_aires") ~ "Americas",
                            community %in% c("india", "beijing", "turkey") ~ "Asia", 
                            community == "auckland" ~ "Oceania"), 
      continent = factor(continent, levels = c("Americas", "Africa", "Europe", "Asia", "Oceania"), ordered = T), 
      community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)", 
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)"), 
      imprecision = abs(clickdistfromtargetcenterx)/160,
      centrality = as.character(abs(5-targetposition)), 
      age_group = floor(ageinyears)
  )

# Filter data set for test trials without audio instructions
data <- all_trials %>%
  filter(trialtype == "test", voiceover == F) %>%
  mutate(trialnr = as.numeric(factor(trialnr))) 
```

# Sample

```{r sample}
# Original sample, check for less than 4 test trials
all_trials %>%
  group_by(subjid) %>%
  summarise(n= n()) %>%
  arrange(n)

# Filter: exclude children with less than 4 test trials (1 training 1, 2 training 2, 1 test with audio instructions + 4 test trials = 8 min)
exclude <- all_trials %>%
  group_by(subjid) %>%
  summarise(n = n()) %>%
  filter(n < 8) %>%
  pull(subjid)

all_trials <- all_trials %>% filter(!subjid %in% exclude)
data <- data %>% filter(!subjid %in% exclude)
```

```{r samplesize_table}
# Create sample size table
sample_table <- data %>%
  group_by(community, age_group) %>%
  summarise(n = n_distinct(subjid)) %>%
  arrange(age_group) %>% 
  pivot_wider(names_from = age_group, values_from = n) %>% 
  arrange(community) %>%
  rename(Community = community) %>% 
  replace(is.na(.), 0) %>% 
  # total per community added as column
  mutate(Total = sum(across(where(is.numeric)))) %>% 
  ungroup() %>% 
  # total per age added as row
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))

saveRDS(sample_table, "../saves/sample_table.rds")
```

# Variation

```{r trialtypes_model}
# Model for variation in imprecision by trial type
# First, prepare data
mdata <- all_trials %>% 
  summarise(
    imprecision = mean(imprecision), 
    .by = c(trialtype, subjid, community)
  ) %>% 
  mutate(
    trialtype = recode(trialtype, touch = "Training 1", fam = "Training 2", test = "Test") %>% as_factor(.) %>% relevel("Test"),
    community = as.factor(community),
  )

# Run model
mtrialtypes <- brm(imprecision ~ trialtype + (trialtype | community),
                    data = mdata,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4) %>%
  add_criterion(c("loo","waic"))

summary(mtrialtypes)

# Extract estimates
# ranef(mtrialtypes)
mestimates <- coef(mtrialtypes)

trialtypes_est <- tibble(
  community = rownames(mestimates$community[,,1]),
  Intercept = mestimates$community[, , "Intercept"][1:17],
  Intercept_q25 = mestimates$community[, , "Intercept"][35:51],
  Intercept_q975 = mestimates$community[, , "Intercept"][52:68],
  Training1 = mestimates$community[, , "trialtypeTraining1"][1:17],
  Training1_q25 = mestimates$community[, , "trialtypeTraining1"][35:51],
  Training1_q975 = mestimates$community[, , "trialtypeTraining1"][52:68],
  Training2 = mestimates$community[, , "trialtypeTraining2"][1:17],
  Training2_q25 = mestimates$community[, , "trialtypeTraining2"][35:51],
  Training2_q975 = mestimates$community[, , "trialtypeTraining2"][52:68]
) %>% 
  mutate(across(where(is.numeric), ~round(., 2)))
trialtypes_est

saveRDS(mtrialtypes, "../saves/mtrialtypes.rds")
# mtrialtypes <- readRDS("../saves/mtrialtypes.rds")
saveRDS(trialtypes_est, "../saves/trialtypes_est")
```

```{r model_draws}
# Extract draws from model for plotting

# newdata <- mdata %>% 
#   select(-imprecision)
# 
# post <- fitted(mtrialtypes, newdata = newdata, re_formula = ~ (trialtype | community)) %>%
#   as_tibble() %>%
#   bind_cols(newdata)
# 
# post_sum <- post %>% 
#   summarise(
#     estimate = mean(Estimate), 
#     .by = c(community, trialtype)
#   ) %>% 
#   pivot_wider(names_from = trialtype, values_from = estimate)
# 
# min(post_sum$`Training 1`)
# min(post_sum$`Training 2`)
```

```{r training_trials}
# Plot for variation in imprecision by trial type
# function to suppress labels
delete_no_display <- function(v) {
  if_else(str_detect(v, 'no_display'), '', v)
}

# prepare plot data
descr_trialtypes <- all_trials %>% 
  mutate(trialtype = recode(trialtype, touch = "Training 1", fam = "Training 2", test = "Test") %>% as_factor(.)) %>%
  na.omit() %>%
  summarise(
    mean_click = mean(imprecision), 
    sd_click = sd(imprecision), 
    .by = trialtype
  )

saveRDS(descr_trialtypes, "../saves/descr_trialtypes.rds")

# plot
training_trials_plot <- all_trials %>% 
  mutate(trialtype = recode(trialtype, touch = "Training 1", fam = "Training 2", test = "Test") %>% as_factor(.)) %>%
  na.omit() %>%
  summarise(mean_click = mean(imprecision), .by = c(community, subjid, ageinyears, trialtype)) %>% 
  mutate(
    # so that first facet row doesnt show x axis labels
    trialtype = if_else(community %in% c("Ocuilan (Mexico)", "Plymouth (UK)", "Pune (India)", "Stanford (USA)", "Windhoek (Namibia)"),
                              as.character(trialtype),
                              paste0(as.character(trialtype), 'no_display')) %>% as_factor(.) %>% fct_relevel(., "Training 1", "Training 2", "Test")
  ) %>%
  ggplot(aes(x = as.factor(trialtype), y = mean_click)) +
  geom_jitter(aes(col = ageinyears), height = 0, width = 0.2, size = 1, alpha = 1) + 
  geom_violin(alpha = 0.25, fill = "darkgrey", scale = "width", draw_quantiles = c(0.5)) +
  scale_color_viridis_c(direction = -1) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, 2)) +
  facet_wrap(~ community, nrow = 3, scales="free_x") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1), 
    legend.position = c(0.93, 0.14), 
    axis.ticks.x = element_blank(),
    strip.text.x = element_text(size = 10), 
    legend.title = element_text(size = 10),
  ) +
  scale_x_discrete(label = delete_no_display) +
  labs(x = "", y = "Average absolute imprecision in target width", col = "Age in years")

training_trials_plot
saveRDS(training_trials_plot, "../saves/training_trials_plot.rds")

ggsave("../figures/tango-cc-variability.pdf", width = 15, height = 11, scale = 0.7, bg = "white")
ggsave("../figures/tango-cc-variability.png", width = 15, height = 11, scale = 0.7, bg = "white")
```

```{r icc}
# Calculate ICC, first prepare data
within_person <- data %>%
  group_by(subjid) %>%
  summarise(
    mean = mean(imprecision, na.rm = TRUE),
    sd = sd(imprecision, na.rm = TRUE),
    var = var(imprecision, na.rm =TRUE),
  )

within_person
saveRDS(within_person, "../saves/within_person.rds")

# proportion of between-community variance relative to the sum of within-community and between-community variance
# generally: ICC = (variance of interest) / (total variance) = (variance of interest) / (variance of interest + unwanted variance)
var_df <- data %>%
  group_by(subjid, community) %>%
  summarise(
    person_mean = mean(imprecision, na.rm = TRUE),
  ) %>% 
  group_by(community) %>%
  summarise(
    community_mean = mean(person_mean),
    community_sd = sd(person_mean),
    within_community_var = var(person_mean)
  ) %>% 
  mutate(
    between_community_var= var(community_mean),
    total_var = between_community_var + sum(within_community_var),
    ICC_b= between_community_var / total_var, 
    ICC_w = sum(within_community_var) / total_var
  ) %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

var_df
saveRDS(var_df, "../saves/var_df.rds")
```

# Internal Consistency
## Odd-even

```{r odd_even_plot}
# Calculate odd-even split
int_con <- data %>%
  mutate(half = ifelse(trialnr %% 2 == 0, "half_1", "half_2")) %>%
  group_by(subjid, community, half) %>%
  summarise(mean_imprecision = mean(imprecision)) %>%
  pivot_wider(names_from = half, values_from = mean_imprecision) %>%
  group_by(community) %>%
  summarise(odd_even = cor(half_1, half_2))

# Plot
odd_even_plot <- int_con %>% 
  left_join(data %>% group_by(community) %>% summarise(n= n_distinct(subjid))) %>% 
  ggplot(aes(x = odd_even, y = fct_rev(community))) + 
  geom_point(aes(size = n), col = "#29485D", alpha = 1) + 
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_size(limits = c(20, 160), breaks = seq(25, 150, 50)) + 
  labs(x = expression(paste("Pearson's ", italic("r"))), y = "", title = "Odd-even split") +
  theme(
    legend.position = c(0.2, 0.2),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0.4, "cm"),
    legend.key.width = unit(-1.2, "cm"),
    title = element_text(size = 8),
    axis.title.x = element_text(size = 8)
  ) +
  expand_limits(y = c(1, 18.5))

odd_even_plot
saveRDS(odd_even_plot, "../saves/odd_even_plot.rds")
```

## Stratified by target position (categorical target centrality)

```{r stratified_data}
# Calculate stratified internal consistency
fn_mean <- function(ds) {
  return(mean(abs(ds$clickdistfromtargetcenterx), na.rm = TRUE))
}

# empty tibble to store results
strat <- tibble()

# loop over communities, 50 iterations, stratified by target position
communities <- unique(data$community)
for (j in c(1:50)) {
  for (i in communities) {
    # stratified by target position
    strat_s <- by_split(
      data = data %>% filter(community == i),
      participants = data %>% filter(community == i) %>% pull(subjid),
      fn_score = fn_mean,
      verbose = F,
      stratification = paste(data %>% filter(community == i) %>% pull(centrality)),
      careful = FALSE
      )

    samp <- tibble(
      community = i,
      iter = j,
      reli = split_coefs(strat_s, cor),
      lci = split_ci(strat_s, cor)$lims["0.025","bca"],
      uci = split_ci(strat_s, cor)$lims["0.975","bca"]
      )

    print(i)
    print(j)
    print(strat_s)
    print(samp)

    strat <- bind_rows(strat, samp)
  }
}

saveRDS(strat, "../saves/stratified_internal_consistency.rds")
# strat <- readRDS("../saves/stratified_internal_consistency.rds")
```

```{r stratified_plot}
# Plot stratified internal consistency
stratified_plot <- strat %>%   
  ggplot(., aes(x = reli, y = fct_rev(community))) +
  geom_density_ridges(col = "white", alpha = 1, rel_min_height = 0.01, fill="#29485D") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = expression(paste("Pearson's ", italic("r"))), y = "", title = "Stratified by target centrality") +
  theme(axis.title.x = element_text(size = 8), title = element_text(size = 8),) +
  expand_limits(y = c(1, 18.5))

stratified_plot
saveRDS(stratified_plot, "../saves/stratified_plot.rds")
```

## Stratified and age-corrected

```{r stratified_agecorrected_data}
# Estimate age-corrected, stratified reliability
# run model first for one community to use update()
split <- split_df(
  data = data %>% filter(community == "Leipzig (Germany)"),
  stratification = data %>% filter(community == "Leipzig (Germany)") %>% pull(centrality),
  careful = FALSE
)

# prepare data
sdata <- bind_rows(
  split[[1]] %>%
    select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
    mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
  split[[2]] %>%
    select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
    mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2")) %>%
  mutate(age = scale(ageinyears))

# run model
strat_age_model <- brm(impr ~ age + (0 + half | subjid),
                data   = sdata,
                family = lognormal(),
                warmup = 1000,
                iter   = 3000,
                control = list(adapt_delta = 0.95),
                chains = 4,
                cores  = 4)

# loop over communities
communities <- unique(data$community)

# empty tibble to store results
strat_age <- tibble()

# loop over communities, 50 iterations, stratified by target position
for (j in c(1:50)) {
  for (i in communities) {

    com_split <- split_df(
      data = data %>% filter(community == i),
      stratification = data %>% filter(community == i) %>% pull(centrality),
      careful = FALSE
    )

    com_sdata <- bind_rows(
      com_split[[1]] %>%
        select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
        mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
      com_split[[2]] %>%
        select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
        mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2")) %>%
      mutate(age = scale(ageinyears))

    # update model for each community
    com_mod <- update(strat_split_model,
        newdata = com_sdata,
        warmup = 1000,
        iter   = 3000,
        control = list(adapt_delta = 0.95),
        chains = 4,
        cores  = 4)

    # extract correlation between testhalves for each community as reliability estimate 
    cor <- summary(com_mod)$random$subjid %>%
      as_tibble(rownames = "parameter") %>%
      filter(parameter == "cor(halfhalf_1, halfhalf_2)") %>%
      mutate(community = i, iter = j)

    strat_age <- bind_rows(strat_age, cor)
  }
}

saveRDS(strat_age, "../saves/age_corrected_internal_consistency.rds")

# strat_age <- readRDS("../saves/age_corrected_internal_consistency.rds")
```

```{r stratified_agecorrected_plot}
# Plot stratified and age-corrected internal consistency
stratified_agecorrected_plot <- strat_age %>% 
  ggplot(., aes(x = Estimate, y = fct_rev(community))) +
  geom_density_ridges(col = "white", alpha = 1, rel_min_height = 0.01, fill="#29485D") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    labs(x = expression(paste("Pearson's ", italic("r"))), y = "", title = "Stratified, age-corrected") +
  theme(axis.title.x = element_text(size = 8), title = element_text(size = 8)) +
  expand_limits(y = c(1, 18.5))

stratified_agecorrected_plot
saveRDS(stratified_agecorrected_plot, "../saves/stratified_agecorrected_plot.rds")
```

```{r combine_data}
# Combine internal consistency data
int_con <- full_join(
  int_con, 
  strat %>% 
    group_by(community) %>%
    summarise(stratified = mean(reli, na.rm = T))
) %>% 
  full_join(
    strat_age %>% 
      group_by(community) %>%
      summarise(stratified_agecorrected = mean(Estimate, na.rm = T))
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

int_con
saveRDS(int_con, "../saves/internal_consistency.rds")
```

# Combine plots

```{r combine_plot}
# Combine plots and arrange for publication figure
odd_even_plot <- readRDS("../saves/odd_even_plot.rds")
stratified_plot <- readRDS("../saves/stratified_plot.rds")
stratified_agecorrected_plot <- readRDS("../saves/stratified_agecorrected_plot.rds")
training_trials_plot <- readRDS("../saves/training_trials_plot.rds")

top_row <- ggarrange(
  odd_even_plot,
  stratified_plot + theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()), 
  stratified_agecorrected_plot + theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()), 
  widths = c(1.7, 1.2, 1.2), 
  nrow = 1
)

ggsave("../figures/tango-cc-reli.pdf", width = 12, height = 6, scale = 0.8, bg = "white")
ggsave("../figures/tango-cc-reli.png", width = 12, height = 6, scale = 0.8, bg = "white")

# top_row <- ggarrange(
#   odd_even_plot,
#   stratified_plot + theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()), 
#   stratified_agecorrected_plot + theme(axis.text.y = element_blank(), axis.ticks.y=element_blank()), 
#   widths = c(1.7, 1.2, 1.2), 
#   nrow = 1,
#   labels = c("B", "C", "", "") 
# )
# 
# ggarrange(
#   training_trials_plot,
#   top_row,
#   nrow = 2,
#   heights = c(3, 1.5),
#   labels = c("A", "") 
# )
# 
# ggsave("../figures/tango-cc-results.png", width = 12, height = 5, scale = 0.7, bg = "white")
```

# Exploratory: Correlation SD and internal consistency

```{r sd_int_plot}
# Calculate correlation between community SD and internal consistency
# general trend: higher internal consistency estimates when larger individual variation

# var_df <- readRDS("../saves/var_df.rds")
# int_con <- readRDS("../saves/internal_consistency.rds")

sd_int_plot <- var_df %>%  
  select(community, community_sd) %>% 
  full_join(int_con) %>% 
  ggplot(aes(x = community_sd, y = stratified_agecorrected)) +
  geom_point(shape = 4, size = 2) +
  geom_smooth(method = "lm", se = FALSE, col = "#29485D") +
  stat_cor(aes(label = paste(..r.label..)), method = "pearson", size = 4, r.accuracy = 0.01, cor.coef.name = "r", label.x = 0.25, label.y = 0.95) +
  scale_x_continuous(limits = c(0.25, 2), breaks = seq(0, 2, 0.5)) +
  scale_y_continuous(limits = c(0.25, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = "Community SD", y = "Internal consistency")

sd_int_plot

sd_int_cor <- var_df %>%  
  select(community, community_sd) %>% 
  full_join(int_con) %>% 
  select(c(community_sd, stratified_agecorrected)) %>%
  corr.test()

sd_int_cor$ci %>% round(2) %>% 
  saveRDS("../saves/sd_int_cor.rds")
```